---
title: "ACST886 Assignment 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Question 1
It is known that the principal amount is $600,000, the rate of interest is 3.5% p.a., and the period is 30 years. The question is asked to calculate the annual instalment.

The formula below is used, where $x$ is the annual instalment, $i$ is the interest rate and $n$ is the number of terms. The uniroot function in r is used to calculate the annual instalment.$$pv=\sum_{k=1}^{n}\frac{x}{(1+i)^n}$$

```{r}
# create a function in terms of instalment to calculate present value
pv <- function(x){
  # create an instalment vector
  instalment <- c(rep(x,30))
  # create an discounting vector
  disc_vec <- 1/(1+0.035)^(1:30)
  # present value is the sum of multiple of two vectors
  pv <- sum(instalment*disc_vec)
  # we want to use the uniroot function
  # hence want to return a value that is zero
  return(pv-600000)
}

## use uniroot to calculate annual instalment
uniroot(pv,c(30000,40000))
```
Hence, the calcuated annual instalment is $32622.80.