---
title: "ACST886 Assignment 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 2
#### (a)
This question is the same as question 1 above.

```{r}
# create a function in terms of instalment to calculate present value
pv <- function(x){
  # create an instalment vector
  instalment <- c(rep(x,30))
  # create an discounting vector
  disc_vec <- 1/(1+0.035)^(1:30)
  # present value is the sum of multiple of two vectors
  pv <- sum(instalment*disc_vec)
  # we want to use the uniroot function
  # hence want to return a value that is zero
  return(pv-600000)
}

## use uniroot to calculate annual instalment
uniroot(pv,c(30000,40000))
```
Hence, the repayment amount is $32,622.8.

#### (b)
For this question, we have total annual payment of \$36,122.8 which is the sum of the annual instalment calculated from part a and the annual interest from the offset account. Since the total annual payment is larger, it means that we can repay the whole loan in less than 30 years. 

In order to calculate the number of terms, the following formula is used. $$600000=\sum_{k=1}^{n}\frac{32622.8+3500}{(1+0.035)^{n}}+\frac{100000}{(1+0.035)^n} $$

Once the number of terms has been calculated, we can calculate amount of the last repayment, which is believed to be less than annual instalment amount.

```{r}
# create a function in terms of instalment to calculate present value
pv <- function(n){
  
  annuity <- function(i,n){
    ## i = interest rate
    ## n = no. of terms
    an <- (1/i)*(1-(1/(1+i)^n))
    return(an)
  }
  
  pv <- (32622.8+3500)*annuity(0.035,n)+100000/(1.035)^n
  return(pv-600000)
}
uniroot(pv,c(0,30))

## we know that the loan can be repaied between 22nd and 23rd year
## hence, the last payment is in year 23

## outstanding loan at the end of year 22
s22 <- ((1+0.035)^22-1)/0.035
loan_22 <- 600000*(1+0.035)^22-(32622.8+3500)*s22
## oustandinf loan at the end of year 23
loan_23 <- loan_22*(1.035)
## At this time, the loan can be fully repaid
## hence to calcualte the adjusted final repayment
last_repayment <- loan_23-3500-100000
last_repayment
```
Hence, the final repayment amount is \$11,484.86.
