---
title: "ACST886 Assignment 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 4

#### Life A

```{r}
require(timevis)
require(lubridate)

## calculate the age for Life A as on the birthday the same year as death
## that is the age as on 16/05/1994
calc_age <- function(birthDate, refDate = Sys.Date()) {

    period <- as.period(new_interval(birthDate, refDate),
                        unit = "year")
    period$year
}
calc_age("1965-05-16","1994-05-16")

## create a timeline for Life A
data1 <- data.frame(
  id      = 1:8,
  content = c("28_years_old","PA_1","1_Jan preceeding Death","29_years_old","Death",
              "PA_2","30_years_old","PA_3"),
  start   = c("1993-05-16","1993-07-21","1994-01-01","1994-05-16","1994-06-22",
              "1994-07-21","1995-05-16","1995-07-21")
)

timevis(data1)
```

As on 16/05/1994, Life A is 29 years old.

(a)
The death of Life A occurs on 22/06/1994 which is around one month after Life A's 29th birthday. Hence, the recorded age at death for the definition age next birthday at death is 30.

(b)
The death of Life A occurs on 22/06/1994 and the 1 January preceeding death is the 1/1/1994. Hence, age last birthday on 1/1/1994 is 28.

(c) 
The death of Life A occurs on 22/06/1994 and it is between the PA_1 and PA_2. The birthday between this peiod is Life A's 29th birthday. Hence, the recoreded age is 29.

#### Life B

```{r}
## calculate the age for Life B as on the birthday the same year as death
## that is the age as on 01/03/1993
calc_age <- function(birthDate, refDate = Sys.Date()) {
  
    period <- as.period(new_interval(birthDate, refDate),
                        unit = "year")
    period$year
}
calc_age("1922-03-01","1993-03-01")

## create a timeline for Life B
data2 <- data.frame(
  id      = 1:8,
  content = c("70_years_old","PA_1","1 Jan preceeding Death","71_years_old","Death",
              "PA_2","72_years_old","PA_3"),
  start   = c("1992-03-01","1992-09-23","1993-01-01","1993-03-01","1993-07-21",
              "1993-09-23","1994-03-01","1994-09-23")
)

timevis(data2)
```

As on 01/03/1993, Life B is 71 years old.

(a)
The death of Life B occurs on 21/07/1993. And the next birthday after death is Life B's 72th birthday. Hence, the recorded age is 72.

(b)
The death of Life B occurs on 21/07/1993 and the 1 January preceeding death is 1/1/1993. Age last birthday preceeding 1/1/1993 is Life B's 70th birthday, Hence, the recoreded age is 70.

(c)
The death of Life B occurs on 21/07/1993 which is between PA_1 and PA_2. The birthday in this period is Life B's 71st birthday. Hence, the recoreded age is 71.


#### Life C

```{r}
## calculate the age for Life C as on the birthday the same year as death
## that is the age as on 21/09/1996
calc_age <- function(birthDate, refDate = Sys.Date()) {

    period <- as.period(new_interval(birthDate, refDate),
                        unit = "year")
    period$year

}
calc_age("1942-09-21","1996-09-21")

## create a timeline for Life C
data3 <- data.frame(
  id      = 1:8,
  content = c("PA_1","53_years_old","1 Jan preceeding Death","Death","PA_2",
              "54_years_old","PA_3","55_years_old"),
  start   = c("1995-06-02","1995-09-21","1996-01-01","1996-04-29","1996-06-02",
              "1996-09-21","1997-06-02","1997-09-21")
)

timevis(data3)
```

As on 21/09/1996, Life C is 54 years old. 

(a)
The death of Life C occurs on 29/04/1996. The next birthday after death is Life C's 54th birthday. Hence, the recoreded age is 54.

(b)
The death of Life C occurs on 29/04/1996 and the 1 January preceeding death is 1/1/1996. Age last birthday preceeding 1/1/1996 is Life C's 53rd birthday. Hence, the recorded age is 53.

(c)
The death of Life C occurs on 29/04/1996 which is between PA_1 and PA_2. The birthday in this peiod is Life C's 53rd birthday. Hence, the recorded age is 53.